---
title: "The reflection of the Syrian civil war on newspaper headlines between 2011-2023 : The Guardian example"
date: "2023-04-21"
output: 
    html_document:
      toc: true
      toc_float: true
      theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("httr")
library("stringr")
library("jsonlite")
library("dplyr")
library("lubridate") #working with dates
library("ggplot2")
library("plotly")

#Store API key password
guardian_api_key <- rstudioapi::askForPassword()

# Create empty vectors to store publication dates and headlines
headline_vector <- vector()
date_vector <- vector()

for (j in 1:500){# loop through 500 pages of resulst to get all relevant headlines
  if (j%%12==0){Sys.sleep(1)}# Slow down request rate i line with API limit
  
  # Download all headlines containing search term "Syria" from 2011-present
  guardian_url_string <- str_c("https://content.guardianapis.com/search?page-size=50&page=", 
  j,
  "&from-date=2011-01-01&q=%22Syria%22&api-key=00198df8-cef5-4a18-b105-c5884070ff53")
  response <- httr::GET(guardian_url_string, query=list(q= "Syria", apiKey=guardian_api_key))
  content <- httr::content(response, as="parse")
  
    # Loop through all 50 entries per page extracting the publication date and headline
    for (i in 1:50){
      current_headline <- content$response$results[[i]]$webTitle[[1]]
      current_date <- content$response$results[[i]]$webPublicationDate
      headline_vector[50*(j-1)+i] <- current_headline
      date_vector[50*(j-1)+i] <- current_date
    }
}

# Export resulting vectors as R objects
saveRDS(date_vector, file = "syria_dates.rds")
saveRDS(headline_vector, file = "syria_headlines.rds")
```

```{r}
##Data Cleaning
### --- Group headlines by month---

# Load vectors containing publication dates and headlines
dates_vector <- readRDS(file = "C:/Users/User/Desktop/data mining/data mining/Guardian-API-Project/Data files/syria_dates.rds")
headline_vector <- readRDS(file = "C:/Users/User/Desktop/data mining/data mining/Guardian-API-Project/Data files/syria_headlines.rds")

# Convert character column to datetime
datetime_vector <- lubridate::ymd_hms(dates_vector)

#Convert vectors to dataframe
headlines_dates <- data.frame(datetime_vector,headline_vector)

#Count headlines per date
headline_month <- headlines_dates %>% 
  mutate(Date=floor_date(as_date(datetime_vector), unit="month")) %>% 
  count(Date)

#Export dataframe containing number of headlines per date
saveRDS(headlines_dates, file = "C:/Users/User/Desktop/data mining/data mining/Guardian-API-Project/Data files/syrian_headlines_dataframe.rds")
```
# Syrian Civil War from 2011-2023 in Guardian Headlines 
## Introduction

The aim of the project is to observe the change and the reflections of the war in the world by analysing the Guardian news headlines from the beginning of the Syrian civil war to the present day (2011-2023). The data set is taken from the Guardian. In total 23,617 headlines are found for the 12 years period. Some of the articles were related to different topics such as earthquake but majority of them related to refugees and certain attacks in Europe. 

### Research Question
- How the headlines related to Syrian Civil War has changed throughout the time (2011-2023)?

### Hypotheses
- While there was a softer discourse in the first couple of years of the war, the discourse changed after 2015 in the opposite direction.
- In particular, the unidentified "terrorist" attacks in Europe in recent years have been associated with the Syrian war and refugees.

### 2. Data Visualisation 1: Headlines by the dates (01-01-2011 to 15-04-2023)

```{r}

#Colorful bar chart

# Group headlines by year
headlines_yearly <- headlines_dates %>%
  mutate(Year = year(datetime_vector)) %>%
  group_by(Year) %>%
  summarise(Headline_Count = n())

# Set custom colors for each year
year_colors <- c("#0072B2", "#E69F00", "#009E73", "#F0E442", "#CC79A7", "#D55E00", "#56B4E9", "#999999", "#FF9DA7", "#D2F53C", "#D25A5A", "#FDAE61", "#AEC7E8", "#FFC0CB", "#B3DE69", "#FB8072", "#80B1D3")

# Create the bar chart
headline_plot <- ggplot(data = headlines_yearly, aes(x = as.factor(Year), y = Headline_Count, fill = as.factor(Year))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = year_colors) +
  ggtitle("Distribution of Headlines on Syria by Year") +
  ylab("Number of Headlines") +
  xlab("Year") +
  theme_minimal()

# Show the plot
headline_plot

#2
#Count headlines per date
headline_month <- headlines_dates %>% 
  mutate(Year = year(datetime_vector),
         Month = month(datetime_vector, label = TRUE)) %>% 
  count(Year, Month, sort = FALSE)

# Define a color palette for each month
month_palette <- c(
  "#FFC107", "#9C27B0", "#E91E63", "#00BCD4", "#8BC34A", "#FF5722",
           "#FF9800", "#4CAF50", "#2196F3", "#3F51B5", "#795548", "#9E9E9E"
)

# Create a vector to store the color of each month
month_colors <- vector(mode = "character", length = 12)

# Assign a color to each month
for (i in 1:12) {
  month_colors[i] <- month_palette[i]
}

# Plot headlines by month and year
headline_plot2 <- ggplot(data = headline_month, aes(x = Year, y = n, fill = Month)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = month_colors) +
  ggtitle("Distribution of Headlines on Syria by Month and Year") +
  ylab("Number of Headlines") +
  xlab("Year") +
  theme_minimal() +
  theme(legend.position = "bottom")
# Show the plot
headline_plot2

```


### 2. Data Visualisation 3. 
```{r}
```



## Results & Conclusion
